---
title: "Figure 1 Analysis"
output: html_document
---
## Manhattan Plots
### Load datasets using CuratedPCa and add fusion info to TCGA
Load mae objects
```{r, message = FALSE}
library(curatedPCaData)
TCGA <- curatedPCaData::mae_tcga
Abida <- curatedPCaData::mae_abida
Baca <- curatedPCaData::mae_baca
Taylor <- curatedPCaData::mae_taylor
Barbieri <- curatedPCaData::mae_barbieri
```
Add fusion status to TCGA mae
```{r}
library(readxl)
TCGA.annotations <- read_xlsx("TCGA sample annotations.xlsx")
for (i in 1:333) {
  index <- which(TCGA@colData@listData$sample_name[i] == TCGA.annotations$SAMPLE_ID)
  TCGA@colData@listData$ERG_fusion_CNA[i] <- TCGA.annotations$Subtype[index]
}
```
### Subset TCGA and Abida datasets by ETS-family fusion status
Subset CNA profiles by fusion status
```{r}
#TCGA subset
TCGA.CNA <- as.data.frame(TCGA@ExperimentList@listData$cna.gistic)
TCGA.CNA.fusion.pos <- TCGA.CNA[,(TCGA@colData@listData$ERG_fusion_CNA == "1-ERG" |
                                    TCGA@colData@listData$ERG_fusion_CNA == "2-ETV1" |
                                    TCGA@colData@listData$ERG_fusion_CNA == "3-ETV4" |
                                    TCGA@colData@listData$ERG_fusion_CNA =="4-FLI1")]
TCGA.CNA.fusion.neg <- TCGA.CNA[,(TCGA@colData@listData$ERG_fusion_CNA == "5-SPOP" |
                                    TCGA@colData@listData$ERG_fusion_CNA == "6-FOXA1" |
                                    TCGA@colData@listData$ERG_fusion_CNA == "7-IDH1" |
                                    TCGA@colData@listData$ERG_fusion_CNA =="8-other")]
#Abida subset
Abida.CNA <- as.data.frame(Abida@ExperimentList@listData$cna.gistic)
Abida.CNA.fusion.pos <- Abida.CNA[,Abida@colData@listData$ERG_fusion_GEX==1]
Abida.CNA.fusion.neg <- Abida.CNA[,Abida@colData@listData$ERG_fusion_GEX==0]
```
Format CNA profiles for enrichment analysis
```{r}
#Format TCGA
TCGA.CNA.fusion.pos <- cbind(rownames(TCGA.CNA), TCGA.CNA.fusion.pos)
colnames(TCGA.CNA.fusion.pos)[1] <- "Hugo_Symbol"
TCGA.CNA.fusion.neg <- cbind(rownames(TCGA.CNA), TCGA.CNA.fusion.neg)
colnames(TCGA.CNA.fusion.neg)[1] <- "Hugo_Symbol"
#Format Abida
Abida.CNA.fusion.pos <- cbind(rownames(Abida.CNA), Abida.CNA.fusion.pos)
colnames(Abida.CNA.fusion.pos)[1] <- "Hugo_Symbol"
Abida.CNA.fusion.neg <- cbind(rownames(Abida.CNA), Abida.CNA.fusion.neg)
colnames(Abida.CNA.fusion.neg)[1] <- "Hugo_Symbol"
```
Compute CNA enrichment in context of ETS fusions
```{r, message=FALSE}
source("analysis scripts.R")
TCGA.loss.enrichment <- concurrent.loss.data(TCGA.CNA.fusion.neg,TCGA.CNA.fusion.pos)
Abida.loss.enrichment <-concurrent.loss.data(Abida.CNA.fusion.neg,Abida.CNA.fusion.pos)
```
### Manhattan Plots for TCGA
Generate manhattan plot objects (GR object) for TCGA patients
```{r, message=FALSE}
source("manhattan.R")
#----------GR object for pval plot----------
#ERG+ Patients
TCGA.loss.fusion.pos <- data.frame(TCGA.loss.enrichment$Hugo_Symbol[TCGA.loss.enrichment$`normalized subtype loss ratio` < 1],
                               TCGA.loss.enrichment$`loss p val`[TCGA.loss.enrichment$`normalized subtype loss ratio` < 1])
TCGA.loss.fusion.pos.gr.pval <- manhattan.object(TCGA.loss.fusion.pos)
#ERG- Patients
TCGA.loss.fusion.neg <- data.frame(TCGA.loss.enrichment$Hugo_Symbol[TCGA.loss.enrichment$`normalized subtype loss ratio` > 1],
                               TCGA.loss.enrichment$`loss p val`[TCGA.loss.enrichment$`normalized subtype loss ratio` > 1])
TCGA.loss.fusion.neg.gr.pval <- manhattan.object(TCGA.loss.fusion.neg)
#----------GR object for OR plot----------
#ERG+ patients
TCGA.loss.fusion.pos <- data.frame(TCGA.loss.enrichment$Hugo_Symbol[TCGA.loss.enrichment$`normalized subtype loss ratio` < 1],
                               (1/TCGA.loss.enrichment$`odds ratio`[TCGA.loss.enrichment$`normalized subtype loss ratio` < 1]))
TCGA.loss.fusion.pos.gr.OR <- manhattan.object(TCGA.loss.fusion.pos)
#ERG- patients
TCGA.loss.fusion.neg <- data.frame(TCGA.loss.enrichment$Hugo_Symbol[TCGA.loss.enrichment$`normalized subtype loss ratio` > 1],
                               TCGA.loss.enrichment$`odds ratio`[TCGA.loss.enrichment$`normalized subtype loss ratio` > 1])
TCGA.loss.fusion.neg.gr.OR <- manhattan.object(TCGA.loss.fusion.neg)
#----------GR object for normalized loss frequency plot----------
#ERG+ patients
TCGA.loss.fusion.pos <- data.frame(TCGA.loss.enrichment$Hugo_Symbol[TCGA.loss.enrichment$`normalized subtype loss ratio` < 1],
                               (1/TCGA.loss.enrichment$`normalized subtype loss ratio`[TCGA.loss.enrichment$`normalized subtype loss ratio` < 1]))
TCGA.loss.fusion.pos.gr.freq <- manhattan.object(TCGA.loss.fusion.pos)
#ERG- patients
TCGA.loss.fusion.neg <- data.frame(TCGA.loss.enrichment$Hugo_Symbol[TCGA.loss.enrichment$`normalized subtype loss ratio` > 1],
                               TCGA.loss.enrichment$`normalized subtype loss ratio`[TCGA.loss.enrichment$`normalized subtype loss ratio` > 1])
TCGA.loss.fusion.neg.gr.freq <- manhattan.object(TCGA.loss.fusion.neg)
```
Generate manhattan plots for loss frequency, and enrichment odds ratio / pval in TCGA
```{r}
#----------pval plot----------
params <- getDefaultPlotParams(plot.type = 4)
params$leftmargin = 0.15
params$topmargin = 20
params$bottommargin = 30
kp <- plotKaryotype(plot.type=4, cex = 0.85, labels.plotter = NULL, plot.params = params)
kpAddChromosomeNames(kp, srt = 90, cex = 0.78, yoffset = 0)
#Add ERG+ p-val manhattan plot
kp <- kpPlotManhattan(kp, TCGA.loss.fusion.pos.gr.pval, r0=0.5, r1=1, ymax=6, points.col = "dark orange", points.cex = 0.4, suggestiveline = 0)
kpAxis(kp, ymin=0, ymax=6, r0=0.5, r1=1, tick.pos = c(0,3,6))
kpAddLabels(kp, labels = "ERG+ Patients\n-log10(p-val)", srt=90, pos=1, r0=0.5, r1=1, cex=1, label.margin = 0.13)
#Add ERG- p-val manhattan plot
kp <- kpPlotManhattan(kp, TCGA.loss.fusion.neg.gr.pval, r0=0.5, r1=0, ymax=6, points.col = "navy", points.cex = 0.4, suggestiveline = 0)
kpAxis(kp, ymin=0, ymax=6, r0=0.5, r1=0, tick.pos = c(0,3,6))
kpAddLabels(kp, labels = "ERG- Patients\n-log10(p-val)", srt=90, pos=1, r0=0.5, r1=0, cex=1, label.margin = 0.13)
kpAddMainTitle(kp, main = "Copy Number Loss Enrichment: TCGA")
#----------OR plot----------
params <- getDefaultPlotParams(plot.type = 4)
params$leftmargin = 0.15
params$topmargin = 20
params$bottommargin = 30
kp <- plotKaryotype(plot.type=4, cex = 0.85, labels.plotter = NULL, plot.params = params)
kpAddChromosomeNames(kp, srt = 90, cex = 0.78, yoffset = 0)
#Add ERG+ OR manhattan plot
kp <- kpPlotManhattan(kp, TCGA.loss.fusion.pos.gr.OR, r0=0.5, r1=1, ymax=10, points.col = "dark orange", points.cex = 0.4, suggestiveline = 1, logp = FALSE, ymin = 1)
kpAxis(kp, ymin=1, ymax=10, r0=0.5, r1=1, tick.pos = c(1,5,10))
kpAddLabels(kp, labels = "ERG+ Patients\n(Odds Ratio)", srt=90, pos=1, r0=0.5, r1=1, cex=1, label.margin = 0.13)
#Add ERG- OR manhattan plot
kp <- kpPlotManhattan(kp, TCGA.loss.fusion.neg.gr.OR, r0=0.5, r1=0, ymax=10, points.col = "navy", points.cex = 0.4, suggestiveline = 1, logp = FALSE, ymin = 1)
kpAxis(kp, ymin=1, ymax=10, r0=0.5, r1=0, tick.pos = c(1,5,10))
kpAddLabels(kp, labels = "ERG- Patients\n(Odds Ratio)", srt=90, pos=1, r0=0.5, r1=0, cex=1, label.margin = 0.13)
kpAddMainTitle(kp, main = "Copy Number Loss Enrichment: TCGA")
#----------Frequency Plot-----------
params <- getDefaultPlotParams(plot.type = 4)
params$leftmargin = 0.15
params$topmargin = 20
params$bottommargin = 30
kp <- plotKaryotype(plot.type=4, cex = 0.85, labels.plotter = NULL, plot.params = params)
kpAddChromosomeNames(kp, srt = 90, cex = 0.78, yoffset = 0)
#Add ERG+ frequency manhattan plot
kp <- kpPlotManhattan(kp, TCGA.loss.fusion.pos.gr.freq, r0=0.5, r1=1, ymax=8, points.col = "dark orange", points.cex = 0.4, suggestiveline = 1, logp = FALSE, ymin = 1)
kpAxis(kp, ymin=1, ymax=8, r0=0.5, r1=1, tick.pos = c(1,4,8))
kpAddLabels(kp, labels = "ERG+ Patients", srt=90, pos=1, r0=0.5, r1=1, cex=1, label.margin = 0.13)
#Add ERG- frequency manhattan plot
kp <- kpPlotManhattan(kp, TCGA.loss.fusion.neg.gr.freq, r0=0.5, r1=0, ymax=8, points.col = "navy", points.cex = 0.4, suggestiveline = 1, logp = FALSE, ymin = 1)
kpAxis(kp, ymin=1, ymax=8, r0=0.5, r1=0, tick.pos = c(1,4,8))
kpAddLabels(kp, labels = "ERG- Patients", srt=90, pos=1, r0=0.5, r1=0, cex=1, label.margin = 0.13)
kpAddMainTitle(kp, main = "Copy Number Loss Frequency: TCGA")
```
### Manhattan plots for Abida
Generate manhattan plot objects (GR object) for Abida patients
```{r, message = FALSE}
source("manhattan.R")
#----------GR object for pval plot----------
#ERG+ Patients
Abida.loss.fusion.pos <- data.frame(Abida.loss.enrichment$Hugo_Symbol[Abida.loss.enrichment$`normalized subtype loss ratio` < 1],
                               Abida.loss.enrichment$`loss p val`[Abida.loss.enrichment$`normalized subtype loss ratio` < 1])
Abida.loss.fusion.pos.gr.pval <- manhattan.object(Abida.loss.fusion.pos)
#ERG- Patients
Abida.loss.fusion.neg <- data.frame(Abida.loss.enrichment$Hugo_Symbol[Abida.loss.enrichment$`normalized subtype loss ratio` > 1],
                               Abida.loss.enrichment$`loss p val`[Abida.loss.enrichment$`normalized subtype loss ratio` > 1])
Abida.loss.fusion.neg.gr.pval <- manhattan.object(Abida.loss.fusion.neg)
#----------GR object for OR plot----------
#ERG+ patients
Abida.loss.fusion.pos <- data.frame(Abida.loss.enrichment$Hugo_Symbol[Abida.loss.enrichment$`normalized subtype loss ratio` < 1],
                               (1/Abida.loss.enrichment$`odds ratio`[Abida.loss.enrichment$`normalized subtype loss ratio` < 1]))
Abida.loss.fusion.pos.gr.OR <- manhattan.object(Abida.loss.fusion.pos)
#ERG- patients
Abida.loss.fusion.neg <- data.frame(Abida.loss.enrichment$Hugo_Symbol[Abida.loss.enrichment$`normalized subtype loss ratio` > 1],
                               Abida.loss.enrichment$`odds ratio`[Abida.loss.enrichment$`normalized subtype loss ratio` > 1])
Abida.loss.fusion.neg.gr.OR <- manhattan.object(Abida.loss.fusion.neg)
#----------GR object for normalized loss frequency plot----------
#ERG+ patients
Abida.loss.fusion.pos <- data.frame(Abida.loss.enrichment$Hugo_Symbol[Abida.loss.enrichment$`normalized subtype loss ratio` < 1],
                               (1/Abida.loss.enrichment$`normalized subtype loss ratio`[Abida.loss.enrichment$`normalized subtype loss ratio` < 1]))
Abida.loss.fusion.pos.gr.freq <- manhattan.object(Abida.loss.fusion.pos)
#ERG- patients
Abida.loss.fusion.neg <- data.frame(Abida.loss.enrichment$Hugo_Symbol[Abida.loss.enrichment$`normalized subtype loss ratio` > 1],
                               Abida.loss.enrichment$`normalized subtype loss ratio`[Abida.loss.enrichment$`normalized subtype loss ratio` > 1])
Abida.loss.fusion.neg.gr.freq <- manhattan.object(Abida.loss.fusion.neg)
```
Generate manhattan plots for loss frequency, and enrichment odds ratio / pval in Abida
```{r}
#----------pval plot----------
params <- getDefaultPlotParams(plot.type = 4)
params$leftmargin = 0.15
params$topmargin = 20
params$bottommargin = 30
kp <- plotKaryotype(plot.type=4, cex = 0.85, labels.plotter = NULL, plot.params = params)
kpAddChromosomeNames(kp, srt = 90, cex = 0.78, yoffset = 0)
#Add ERG+ p-val manhattan plot
kp <- kpPlotManhattan(kp, Abida.loss.fusion.pos.gr.pval, r0=0.5, r1=1, ymax=6, points.col = "dark orange", points.cex = 0.4, suggestiveline = 0)
kpAxis(kp, ymin=0, ymax=6, r0=0.5, r1=1, tick.pos = c(0,3,6))
kpAddLabels(kp, labels = "ERG+ Patients\n-log10(p-val)", srt=90, pos=1, r0=0.5, r1=1, cex=1, label.margin = 0.13)
#Add ERG- p-val manhattan plot
kp <- kpPlotManhattan(kp, Abida.loss.fusion.neg.gr.pval, r0=0.5, r1=0, ymax=6, points.col = "navy", points.cex = 0.4, suggestiveline = 0)
kpAxis(kp, ymin=0, ymax=6, r0=0.5, r1=0, tick.pos = c(0,3,6))
kpAddLabels(kp, labels = "ERG- Patients\n-log10(p-val)", srt=90, pos=1, r0=0.5, r1=0, cex=1, label.margin = 0.13)
kpAddMainTitle(kp, main = "Copy Number Loss Enrichment: Abida")
#----------OR plot----------
params <- getDefaultPlotParams(plot.type = 4)
params$leftmargin = 0.15
params$topmargin = 20
params$bottommargin = 30
kp <- plotKaryotype(plot.type=4, cex = 0.85, labels.plotter = NULL, plot.params = params)
kpAddChromosomeNames(kp, srt = 90, cex = 0.78, yoffset = 0)
#Add ERG+ OR manhattan plot
kp <- kpPlotManhattan(kp, Abida.loss.fusion.pos.gr.OR, r0=0.5, r1=1, ymax=10, points.col = "dark orange", points.cex = 0.4, suggestiveline = 1, logp = FALSE, ymin = 1)
kpAxis(kp, ymin=1, ymax=10, r0=0.5, r1=1, tick.pos = c(1,5,10))
kpAddLabels(kp, labels = "ERG+ Patients\n(Odds Ratio)", srt=90, pos=1, r0=0.5, r1=1, cex=1, label.margin = 0.13)
#Add ERG- OR manhattan plot
kp <- kpPlotManhattan(kp, Abida.loss.fusion.neg.gr.OR, r0=0.5, r1=0, ymax=10, points.col = "navy", points.cex = 0.4, suggestiveline = 1, logp = FALSE, ymin = 1)
kpAxis(kp, ymin=1, ymax=10, r0=0.5, r1=0, tick.pos = c(1,5,10))
kpAddLabels(kp, labels = "ERG- Patients\n(Odds Ratio)", srt=90, pos=1, r0=0.5, r1=0, cex=1, label.margin = 0.13)
kpAddMainTitle(kp, main = "Copy Number Loss Enrichment: Abida")
#----------Frequency Plot-----------
params <- getDefaultPlotParams(plot.type = 4)
params$leftmargin = 0.15
params$topmargin = 20
params$bottommargin = 30
kp <- plotKaryotype(plot.type=4, cex = 0.85, labels.plotter = NULL, plot.params = params)
kpAddChromosomeNames(kp, srt = 90, cex = 0.78, yoffset = 0)
#Add ERG+ frequency manhattan plot
kp <- kpPlotManhattan(kp, Abida.loss.fusion.pos.gr.freq, r0=0.5, r1=1, ymax=8, points.col = "dark orange", points.cex = 0.4, suggestiveline = 1, logp = FALSE, ymin = 1)
kpAxis(kp, ymin=1, ymax=8, r0=0.5, r1=1, tick.pos = c(1,4,8))
kpAddLabels(kp, labels = "ERG+ Patients", srt=90, pos=1, r0=0.5, r1=1, cex=1, label.margin = 0.13)
#Add ERG- frequency manhattan plot
kp <- kpPlotManhattan(kp, Abida.loss.fusion.neg.gr.freq, r0=0.5, r1=0, ymax=8, points.col = "navy", points.cex = 0.4, suggestiveline = 1, logp = FALSE, ymin = 1)
kpAxis(kp, ymin=1, ymax=8, r0=0.5, r1=0, tick.pos = c(1,4,8))
kpAddLabels(kp, labels = "ERG- Patients", srt=90, pos=1, r0=0.5, r1=0, cex=1, label.margin = 0.13)
kpAddMainTitle(kp, main = "Copy Number Loss Frequency: Abida")
```

## Coexclusivity Analysis
### Load datasets using CuratedPCa and add fusion info to TCGA
Load mae objects
```{r, message = FALSE}
library(curatedPCaData)
TCGA <- curatedPCaData::mae_tcga
Abida <- curatedPCaData::mae_abida
```
Add fusion status to TCGA mae
```{r}
library(readxl)
TCGA.annotations <- read_xlsx("TCGA sample annotations.xlsx")
for (i in 1:333) {
  index <- which(TCGA@colData@listData$sample_name[i] == TCGA.annotations$SAMPLE_ID)
  TCGA@colData@listData$ERG_fusion_CNA[i] <- TCGA.annotations$Subtype[index]
}
```
### Subset TCGA and Abida by fusion status and compute enrichment
Subset CNA profiles by fusion status
```{r}
#TCGA subset
TCGA.CNA <- as.data.frame(TCGA@ExperimentList@listData$cna.gistic)
TCGA.CNA.fusion.pos <- TCGA.CNA[,(TCGA@colData@listData$ERG_fusion_CNA == "1-ERG" |
                                    TCGA@colData@listData$ERG_fusion_CNA == "2-ETV1" |
                                    TCGA@colData@listData$ERG_fusion_CNA == "3-ETV4" |
                                    TCGA@colData@listData$ERG_fusion_CNA =="4-FLI1")]
TCGA.CNA.fusion.neg <- TCGA.CNA[,(TCGA@colData@listData$ERG_fusion_CNA == "5-SPOP" |
                                    TCGA@colData@listData$ERG_fusion_CNA == "6-FOXA1" |
                                    TCGA@colData@listData$ERG_fusion_CNA == "7-IDH1" |
                                    TCGA@colData@listData$ERG_fusion_CNA =="8-other")]
#Abida subset
Abida.CNA <- as.data.frame(Abida@ExperimentList@listData$cna.gistic)
Abida.CNA.fusion.pos <- Abida.CNA[,Abida@colData@listData$ERG_fusion_GEX==1]
Abida.CNA.fusion.neg <- Abida.CNA[,Abida@colData@listData$ERG_fusion_GEX==0]
```
Format CNA profiles for enrichment analysis
```{r}
#Format TCGA
TCGA.CNA.fusion.pos <- cbind(rownames(TCGA.CNA), TCGA.CNA.fusion.pos)
colnames(TCGA.CNA.fusion.pos)[1] <- "Hugo_Symbol"
TCGA.CNA.fusion.neg <- cbind(rownames(TCGA.CNA), TCGA.CNA.fusion.neg)
colnames(TCGA.CNA.fusion.neg)[1] <- "Hugo_Symbol"
#Format Abida
Abida.CNA.fusion.pos <- cbind(rownames(Abida.CNA), Abida.CNA.fusion.pos)
colnames(Abida.CNA.fusion.pos)[1] <- "Hugo_Symbol"
Abida.CNA.fusion.neg <- cbind(rownames(Abida.CNA), Abida.CNA.fusion.neg)
colnames(Abida.CNA.fusion.neg)[1] <- "Hugo_Symbol"
```
Compute CNA enrichment in context of ETS fusions
```{r, message=FALSE}
source("analysis scripts.R")
TCGA.loss.enrichment <- concurrent.loss.data(TCGA.CNA.fusion.neg,TCGA.CNA.fusion.pos)
Abida.loss.enrichment <-concurrent.loss.data(Abida.CNA.fusion.neg,Abida.CNA.fusion.pos)
```
### Compute threshold
threshold <-